---
title: "Strings and Factors"
output: github_document
---

```{r}
library(rvest)
library(tidyverse)
library(p8105.datasets)
```


## Strings and regex

Define string vector and filter specific value based on string variable (case-sentitive)
```{r}
string_vec = c("my", "name", "is", "jeff")

str_detect(string_vec, "jeff")
str_detect(string_vec, "e") 
```

Replace string value and white space
```{r}
str_replace(string_vec, "jeff", "Jeff")
str_replace(string_vec, "e", "")

str_remove(string_vec, "m")
```

Patterns matter when matching 
```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "i think")
str_detect(string_vec, "^i think")  # match with start with ""
str_detect(string_vec, "i think$")  # match with end with ""
``` 

```{r}
str_detect(string_vec, "i think$")
```

`[]` anything inside counts for matching, and anything outside doesn't count
```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )

str_detect(string_vec,"[Pp]umpkin")
```

More complicated cases: number follows letter
```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "^[0-9][a-zA-Z]") # detect any number between 0-9, A-Z, and a-z
```

```{r}
string_vec = c(
  'Its 7:11 in the evening',
  'want to go to 7-11?',
  'my flight is AA711',
  'NetBios: scanning ip 203.167.114.66'
  )

str_detect(string_vec, "7.11") # . is a placeholder for anything between 7 and 11
str_detect(string_vec, "7\\.11") # indicates . is a special character trying to detect
```

Some characters are “special”. These include `[` and `]`, `(` and `)`, and .. If you want to search for these, you have to indicate they’re special using `\`. 
Unfortunately, `\` is also special, so things get weird.
```{r}
string_vec = c(
  'The CI is [2, 5]',
  ':-]',
  ':-[',
  'I found the answer on pages [6-7]'
  )

str_detect(string_vec, "\\[")
```


## Thoughts on factors

Default ordering is alphabatecial
```{r}
vec_sex = factor(c("male", "male", "female", "female"))
vec_sex
```

Force it to be numeric variable
```{r}
as.numeric(vec_sex)
```

```{r}
vec_sex = fct_relevel(vec_sex, "male") 
vec_sex
```

Reorder
```{r}
vec_sex = fct_inorder(vec_sex)
vec_sex
```

```{r}
as.numeric(vec_sex)
```


## NSDUH

Import from web 
```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

table_marj = 
  read_html(nsduh_url) |> 
  html_table() |> 
  first() |>
  slice(-1)
```

Tidy the dataset
```{r}
data_marj = 
  table_marj |>
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |> # what to look for when separate: (
  mutate(
    year = str_replace(year, "\\)", ""), # remove also work, "\\)"
    percent = str_replace(percent, "[a-c]$", ""), # $ specifies only a-c letter
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West"))) # !(State%in% c()) to delete those only
```

Make a quick plot
```{r}
data_marj |>
  filter(age == "12-17") |> 
  mutate(State = fct_reorder(State, percent)) |> # mutate order just for plot, not for the whole dataset, so keep it here instead in the clean step.
  ggplot(aes(x = State, y = percent, color = year)) + 
    geom_point() + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```



## Restaurant inspections


```{r}
data("rest_inspec")

rest_inspec = 
    rest_inspec |> 
    rename(borough = boro)


rest_inspec |> 
  group_by(borough, grade) |> 
  summarize(n = n()) |> 
  pivot_wider(names_from = grade, values_from = n)
```

Only Include A/B/C grade
```{r}
rest_inspec =
  rest_inspec |>
  filter(grade %in% c("A", "B", "C"), borough != "Missing") |> 
  mutate(borough = str_to_title(borough))

rest_inspec
```

Look for Pizza place
```{r}
rest_inspec |> 
  filter(str_detect(dba, "Pizza")) |> 
  group_by(borough, grade) |> 
  summarize(n = n()) |> 
  pivot_wider(names_from = grade, values_from = n)
```

```{r}
rest_inspec |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |> #any possible combination for those letters in the []
  group_by(borough, grade) |> 
  summarize(n = n()) |> 
  pivot_wider(names_from = grade, values_from = n)
```

Make a plot instead
```{r}
rest_inspec |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |>
  ggplot(aes(x = borough, fill = grade)) + 
  geom_bar() 
```

```{r}
rest_inspec |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |>
  mutate(borough = fct_infreq(borough)) |> # order the bars according to frequency
  ggplot(aes(x = borough, fill = grade)) + 
  geom_bar() 
```

```{r}
rest_inspec |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |>
  mutate(
    borough = fct_infreq(borough), #convert string variables to factor
    borough = str_replace(borough, "Manhattan", "The City")) |> # replace names, and convert back to strings, then the order change to alphabetical
  ggplot(aes(x = borough, fill = grade)) + 
  geom_bar() 
```

`replace` to rename.
```{r}
rest_inspec |> 
  filter(str_detect(dba, "[Pp][Ii][Zz][Zz][Aa]")) |>
  mutate(
    borough = fct_infreq(borough),
    borough = replace(borough, which(borough == "Manhattan"), "The City")) |>
  ggplot(aes(x = borough, fill = grade)) + 
  geom_bar()
```
Not work at all. Factors have very specific values, trying to use a value that is not an existing factor level won’t work.
 
Use `fct_recode` to directly rename factor levels
```{r}
rest_inspec |> 
  filter(str_detect(dba, regex("pizza", ignore_case = TRUE))) |>
  mutate(
    borough = fct_infreq(borough),
    borough = fct_recode(borough, "The City" = "Manhattan")) |>
  ggplot(aes(x = borough, fill = grade)) + 
  geom_bar()
```


## Weather Data


```{r}
library(p8105.datasets)
data("weather_df")
```

Reordered name “by hand” using `fct_relevel`:
```{r}
weather_df |>
  mutate(name = forcats::fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))) |> 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```

  Reorder name according to tmax values in each name using `fct_reorder`:
```{r}
weather_df |>
  mutate(name = forcats::fct_reorder(name, tmax)) |> 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```
Ordering of factor variables play an important role for linear regression,
the ordering determines the “reference” category
```{r}
weather_df |>
  lm(tmax ~ name, data = _)
```
Change the reference group by `fct_relevel` for regression
```{r}
weather_df |>
  mutate(name = forcats::fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))) |> 
  lm(tmax ~ name, data = _)
```


## Pulse Data

```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") |>
  janitor::clean_names() |>
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") |>
  select(id, visit, everything()) |>
  mutate(
    visit = str_replace(visit, "bl", "00m"),
    visit = factor(visit)) |>
  arrange(id, visit)

print(pulse_data, n = 12)
```


## Airbnb


```{r}
data("nyc_airbnb")

nyc_airbnb |>
  filter(neighbourhood_group == "Manhattan") |> 
  mutate(
    neighbourhood = fct_reorder(neighbourhood, price)) |> 
  ggplot(aes(x = neighbourhood, y = price)) +
  geom_boxplot() +
  coord_flip() + 
  ylim(0, 1000)
```

